---
title: "den-weather"
format:
  gfm:
    echo: false
    message: false
    warning: false
---

## Denver Weather Patterns
```{r}
#| output: false
suppressPackageStartupMessages(library(tidyverse))
library(nombre)
library(janitor)
source('R/functions.R')
```


## References
- <https://www.ncei.noaa.gov/data/global-summary-of-the-day/doc/readme.pdf>

```{r}
daily <- 
list.files('data/daily/raw', full.names = T) |> 
  map(~read_csv(., show_col_types = F) |> format_daily()) |> 
  bind_rows() |> 
  janitor::clean_names()

dy <- 
daily |> 
  filter(temp_attributes == 24) |> 
  slice_max(date)

comparable <- 
{dy$day_of_year +
  days(-7:7)} |> 
  `year<-`(2000) |> 
  tibble(day_of_year = _) |> 
  left_join(daily, by = 'day_of_year') |> 
  filter(temp_attributes == 24) |> 
  filter(date != dy$date)

last_update <-
  read_lines('now.txt') |> 
    ymd_hms() |> 
    with_tz('America/Denver') |> 
    format('%F %X %Z')
```

# High/Low data for `r dy$date`

## Updated at
`r last_update`

## Highs
```{r}
high_percentile <-
  comparable |>
  summarise(mean(actual_high < dy$actual_high, na.rm = T)) |> 
  pull()

high_percentile_string <- str_glue("This day's high of __{dy$actual_high}F__ was in the __{nom_ord(round(high_percentile*100))} percentile__ for highs in similar days.")

high_hist <- 
comparable |> 
  ggplot(aes(x = actual_high)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = dy$actual_high) +
  ggtitle('This day\'s high vs highs of comparable days')

last_time_this_warm_or_warmer <- 
daily |> 
  filter(temp_attributes == 24,
         date != dy$date,
         actual_high >= dy$actual_high) |> 
  slice_max(date)

high_last_time_string <- str_glue("The last time the high was this high (or higher) was __{last_time_this_warm_or_warmer$date}__, when the high was __{last_time_this_warm_or_warmer$actual_high}F__.")
```

`r high_percentile_string`  
`r high_last_time_string`
```{r}
high_hist
```


## Lows
```{r}
low_percentile <-
  comparable |>
  summarise(mean(actual_low < dy$actual_low, na.rm = T)) |> 
  pull()

low_percentile_string <- str_glue("This day's low of __{dy$actual_low}F__ was in the __{nom_ord(round(low_percentile*100))} percentile__ for lows in similar days.")

low_hist <- 
comparable |> 
  ggplot(aes(x = actual_low)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = dy$actual_low) +
  ggtitle('This day\'s low vs lows of comparable days')

last_time_this_cold_or_colder <- 
daily |> 
  filter(temp_attributes == 24,
         date != dy$date,
         actual_low <= dy$actual_low) |> 
  slice_max(date)

low_last_time_string <- str_glue("The last time the low was this low (or lower) was __{last_time_this_cold_or_colder$date}__, when the high was __{last_time_this_cold_or_colder$actual_low}F__.")
```

`r low_percentile_string`  
`r low_last_time_string`

```{r}
low_hist
```


## TODO

- [x] Format Daily
- [x] Today’s low was in the __th percentile for similar days (+/- 7 days)
- [x] Today’s low was the coldest since ___
- [ ] Today’s low was the warmest since ___
- [ ] Today’s high was the coldest since ___
- [x] Today’s high was the warmest since ___
